Invasive lionfish present region-wide variation of allometric growth in the Western Atlantic

### 2018-06-01 docs/Manuscript.Rmd
Updated manuscript, ready for (last?) round of revision with Sean

---
title: Invasive lionfish present region-wide variation of allometric growth in the
  Western Atlantic
author: "Juan Carlos Villaseñor-Derbez$^1$, Sean Fitzgerald"
header-includes:
- \usepackage{float}
- \floatplacement{figure}{H}
output:
  pdf_document:
    keep_tex: yes
    number_sections: no
    template: peerj_template.tex
keywords: A, B, C
institution: Bren School of Environmental Sciences and Management, University of California
  Santa Barbara, Santa Barbara, California, USA
abstract: "Lionfish (*Pterois volitans/miles*) are an invasive species in the Western Atlantic and the Caribbean. Improving management of invasive lionfish populations requires accurate estimates of total biomass, which in turn depend on accurate definition of the length-weight relationship. Here, we reviewed published length-weight relationships of lionfish taken from throughout their invasive range and found that lionfish of equal lengths have lower body mass in the Caribbean than in the Atlantic or Gulf of Mexico. Additionally, we report a new pair of length-weight parameters ($a = 3.2056 \\times 10^{-6}; b = 3.235$) for organisms sampled in the central Mexican Caribbean region. The substantial spatial variation  in length-weight parameters highlights the importance of using site-specific information when estimating lionfish biomass using length observations. These findings can have major implications for management in terms of predicting effects on local ecosystems, evaluating the effectiveness of removal programs, or estimating biomass available for harvest."
---

@@

```{r load packages}
suppressPackageStartupMessages({
  # library(RColorBrewer)
  # library(colorRamps)
  # library(e1071)
  library(stringr)
  library(agricolae)
  library(magrittr)
  library(stargazer)
  library(sandwich)
  library(lmtest)
  library(broom)
  library(tmap)
  library(rgdal)
  library(sf)
  library(cowplot)
  library(tidyverse)
})
```

@@

Lionfish (*Pterois volitans/miles* complex) are an invasive species in the western Atlantic and the Caribbean, likely introduced through liberation of aquarium-kept organisms \citep{betancurr_2011}. They are the first invasive marine vertebrates established along the North Atlantic Caribbean coasts \citep{schofield_2009,schofield_2010,sabidoitza_2016} and their presence has been labeled as a major marine invasion because they threaten local biodiversity, spread rapidly, and are difficult to manage \citep{hixon_2016}. Invasive lionfish are primarily studied in coral reef ecosystems, where their impacts are far-reaching. For example, field experiments by \citet{albins_2008} showed that lionfish establishment led to reduced recruitment of native fishes by nearly 80% over a five week period in Florida, and increased lionfish biomass coincided with a 65% reduction in the biomass of native prey fishes along Bahamian coral reefs in just two years \citep{green_2012}. Lionfish have also established invasive populations in other habitats such as estuaries, mangroves , hard-bottomed areas , and mesophotic reefs \citep{jud_2011,barbour_2010,muoz_2011,andradibrown_2017,claydon_2012}.

A substantial amount of research describes lionfish feeding ecology from North Carolina to Costa Rica \citep{muoz_2011,morris_2009,cote_2013,dahl_2014,valdezmoreno_2012,villaseorderbez_2014,hackerott_2017,sandel_2015}. A meta-analysis by  \citet{peake_2018} showed that invasive lionfish prey on at least 167 different species across the tropical and temperate North Atlantic. Their feeding behavior and high consumption rates can reduce recruitment and population sizes of native reef-fish species, and can further endanger reef fish \citep{albins_2008, green_2012,rocha_2015}. (However, see \citet{hackerott_2017} for a case where there was no evidence that lionfish affected the density, richness, or community composition of prey fishes). Major efforts have been made to understand the possible impacts of the invasion by tracking the spread of established lionfish populations through time \citep{schofield_2009,schofield_2010} and by predicting invasion ranges under future climates \citep{grieve_2016}. Trophic impacts of lionfish can be minimized if local lionfish biomass is controled by by culling \citep{ariasgonzalez_2011}. However, the feasability of controling local populations remains a topic of debate \citep{cote_2014}.

@@

The feasibility of establishing fisheries through lionfish removal programs has been extensively evaluated through field observations and empirical modeling \citep{barbour_2011,morris_2011,deleon_2013,johnston_2015,sandel_2015,chin_2016,usseglio_2017}. One contributing factor to the success of many removal programs is the sedentary nature of adult lionfish \citep{jud_2012}. Culling programs are effective in reducing adult populations largely because lionfish exhibit high levels of site fidelity and rarely leave their home range in most cases \citep{Fishelson_1997,cote_2014,kochzius_2005} -- But see \citet{andradibrown_2017} for cases when deep-water populations maintain shallow-water populations. Fish with this sedentary behavior are likely to exhibit high levels of spatial variation in important life history characterstics such as growth or natural mortality rate \citep{hutchinson_2008,wilson_2012}. The importance of considering spatial heterogeneity is well-documented in terms of assessing and managing sedentary species \citep{gunderson_2008,guan_2013}, and such variation should be accounted for when evaluating the feasibility of establishing lionfish fisheries as well.

Empirical modeling efforts examining the feasibility of establishing fisheries for lionfish involve modeling changes in biomass in response to changes in mortality (*i.e.* culling). A common way to model this is via length-structured population models, where fish lengths are converted to  weight in order to calculate total biomass \citep{cote_2014,barbour_2011,andradibrown_2017}. The length-weight relationship is therefore an essential component of these models, but this relationship can vary across regions as a response to biotic (e.g. local food availability) and abiotic (e.g. water temperature) conditions \citep{johnson_2016}. Literature suggests that site-specific parameters are necessary in order to accurately estimate biomass when length-weight relationships are spatially variable , and this variability becomes increasingly important when estimating the potential effectiveness of (and resources needed for)  lionfish culling programs or when identifying total biomass available for harvest by fishers \citep{barbour_2011,morris_2011,johnston_2015,chin_2016,cote_2014}. In addition to environmentally-driven spatial variation, genetic analysis of invasive lionfish suggest biological differences due to the existence of two genetically distinct subpopulations between the northwest Atlantic and the Caribbean \citep{betancurr_2011}. To date, no studies have examined region-wide differences in length-weight parameters despite the large number of studies reporting this relationship for lionfish.

@@

We reviewed 12 published studies and obtained 17 length-weight relationships for the North Atlantic (n = 1), Gulf of Mexico (n = 7,), and Caribbean (n = 10, Table \ref{tab:all_params}, Fig \ref{fig:all_allo}). We collected information on sampling methods, sex differentiation, location, and depth ranges from each study when available, and assumed both genders were included in a study if gender was unspecified. Two studies reported parameters for genders combined an separated \citep{aguilarperera_2016,fogg_2013}, while the rest presented pooled results. Reviewed studies presented information for organisms obtained at depths between `r min(review$minD, na.rm = T)` and `r max(review$maxD, na.rm = T)` m. Three studies explicitly stated that their organisms were sampled with pole spears \citep{aguilarperera_2016,chin_2016,dahl_2014,sabidoitz_2016}, and five studies mentioned that some of their organisms were obtained with pole spears (or other type of harpoon) but also hand-held nets or fish traps \citep{sandel_2015,barbour_2011,fogg_2013,edwards_2014,sabidoitza_2016,sabidoitz_2016,toledohernndez_2014}, and two studies did not specify how organisms were sampled \citep{deleon_2013,darling_2011}. \citet{fogg_2013} use spine-less weight in the length-weight relationship estimation, and thus their parameters likely underestimate whole wieght. Since no spine-less to whole weight conversions were available, these parameters were taken as reported.

@@

```{r, fig.cap = "\\label{fig:map}Locations where allometric growth parameters of lionfish (*Pterois spp*) have been reported. Circle sizes indicate sample size from each study, colors indicate the $b$ coefficient from Eq. \\ref{eq:allometric}."}

@@

#Small shapefile for region of interest
xy1 <- data.frame(X = mean(-87.30662, -87.19194), Y = mean(20.39283, 20.51278))
coordinates(xy1) <- c("X", "Y")
proj4string(xy1) <- proj
res1 <- spTransform(xy1, proj) %>% 
  st_as_sf()

#Create shapefile for sampling locations
sample_points <- fish %>%
  group_by(Location) %>%
  summarize(latitude = mean(Latitude, na.rm = T),
            longitude = mean(Longitude, na.rm = T)) %>%
  ungroup()

xy <- data.frame(X = -1*abs(sample_points$longitude), Y = sample_points$latitude)
coordinates(xy) <- c("X", "Y")
proj4string(xy) <- proj  ## for example
res <- spTransform(xy, proj)
b <- SpatialPointsDataFrame(coords = res, data = sample_points, proj4string = proj)

# Create shapefile for relevant communities
Communities <- data.frame(Longitude = c(-87.24),
                           Latitude = c(20.5),
                           Location = c("Pto. Aventuras"),
                           stringsAsFactors = F)

xy2 <- data.frame(X = -1*abs(Communities$Longitude), Y = Communities$Latitude)
coordinates(xy2) <- c("X", "Y")
proj4string(xy2) <- proj  ## for example
res2 <- spTransform(xy2, proj)
c <- SpatialPointsDataFrame(coords = res2, data = Communities, proj4string = proj)

# Map with SF

ggplot() +
  geom_sf(data = W) +
  geom_point(data = review %>%
      mutate(N = ifelse(is.na(N), 1, N)),
    aes(x = lon, y = lat, fill = b, size = N), shape = 21) +
  theme_classic() +
  scale_fill_gradientn(colors = colorRamps::matlab.like(20)) +
  theme(legend.background = element_blank(),
        panel.grid.major = element_line(colour = "transparent")) +
  lims(x = c(-100, -60), y = c(9, 36)) +
  labs(x = "longitude", y = "latitude") +
  geom_text(aes(x = -70, y = 30), label = "North Atlantic\nOcean") +
  geom_text(aes(x = -90, y = 25), label = "Gulf of Mexico") +
  geom_text(aes(x = -75, y = 15), label = "Caribbean\nSea")

@@

<!-- Still having trouble to explain this paragraph. Perhaps you can make it better? The idea is that we test wether the differences are trivial by taking all 109 lionfish and calculating TW with each of the 18 available relationships. We then take those, and divide them by the true weights, and obtain a ratio that is easy to interpret if theres an over or under estimation... andwe do some bassic stats on them -->

Combining the length-weight parameters extracted from the literature and the additional pair calculated here, we obtain a total of 18 pairs. We use the central Mexican Caribbean as a case study of how the use of *ex situ* parameters influences the accuracy of weight estimates for lionfish. We estimated TW from the TL observations we collected in the central Mexican Caribbean (n = 109) using each of the 18 parameter pairs and divided predicted weights by known observed weights to obtain a simple measure of over- or underestimation. Difference in mean weight ratios across the different parameter pairs were tested with a one-way analysis of variance (ANOVA). All analyses were performed in R version `r paste(version$major, version$minor, sep = ".")` \citep{rcore_2018} . Raw data and code used in this work are available at dryad.org.

@@

Inserting TL observations for the central Mexican Caribbean into \ref{eq:log-alo} Eq. 2 and converting to \ref{eq:log-alo-trans} resulted in the coefficient values $b = `r b`$ and $c = `r c`$ ($R^2 = `r Rsq`$, `r Fstring`, $p < 0.001$). The length-weight coefficients estimated in this study were within the range identified by studies in other regions (Table \ref{tab:all_params}). The significant difference between this allometric factor ($b$) and the isometric growth factor $b = 3$ ($t(107) = `r formatC((coef(model)[[2]]-3)/robust_se[[2]], digits = 2, format = "f")`; p<0.001$) indicated that lionfish present allometric growth, which is consistent with other studies. Figure \ref{fig:l-w-carib} shows the relationship between TL and TW for this region, and more information on model fit is presented in Table S2.

Parameters from models fit to males or females exclusively tend to have a higher steepness (*i.e.* higher allometric parameter), with mean $\pm$ standard deviation values of $b = `r formatC(mean(review$b[review$Gender == "Male"]), digits = 2, format = "f")` \pm `r formatC(sd(review$b[review$Gender == "Male"]), digits = 2, format = "f")`$ and $b = `r formatC(mean(review$b[review$Gender == "Female"]), digits = 2, format = "f")` \pm `r formatC(sd(review$b[review$Gender == "Female"]), digits = 2, format = "f")`$ for males and females respectively, compared to parameters from models for pooled genders with a mean $\pm$ standard deviation value of $b = `r formatC(mean(review$b[review$Gender == "Both"]), digits = 2, format = "f")` \pm `r formatC(sd(review$b[review$Gender == "Both"]), digits = 2, format = "f")`$. In the case of the ponderal index ($a$) and its $log_{10}$ transformation ($c$), values were higher for parameters for pooled genders. Figure \ref{fig:all_allo} shows the length-weight relationships with parameters from all studies.

There were significant differences in our predicted weights for the central Mexican Caribbean when using the different pairs of parameters ($F(df = 15; 1728) = 38.26; p < 0.001$). The lowest weight estimates resulted from using the allometric parameters from Banco Chinchorro in the Caribbean \citet{sabidoitz_2016}, and the highest weight estimates came from the Northern Atlantic\citet{barbour_2011}. The calculated ratio of predicted-to-observed weight  ranged from `r filter(Predicted_biomass, Study_id == "Sabido2BC") %$% mean(Biomass_ratio) %>% formatC(digits = 2, format = "f")` $\pm$ `r filter(Predicted_biomass, Study_id == "Sabido2BC") %$% sd(Biomass_ratio) %>% formatC(digits = 2, format = "f")` to `r filter(Predicted_biomass, Study_id == "Barbour") %$% mean(Biomass_ratio) %>% formatC(digits = 2, format = "f")` $\pm$ `r filter(Predicted_biomass, Study_id == "Barbour") %$% sd(Biomass_ratio) %>% formatC(digits = 2, format = "f")` (mean $\pm$ SD). Predicted-to-observed weight ratios are presented in Figure \ref{fig:bio_ratio}. Spine-less weight parameters from \citet{fogg_2013} still produced predicted-to-observed weight rations > 1.

@@

```{r}
review %>% 
  select(Region, Gender2, N, a, b, c, r2, Study) %>% 
  filter(!Study == "Fishbase") %>% 
  arrange(Region) %>% 
  mutate(r2 = ifelse(is.na(r2), "-", r2),
         N = ifelse(is.na(N), "-", N),
         a = formatC((a*1e5), digits = 2)) %>% 
  knitr::kable(format = "latex",
               row.names = F,
               digits =  c(0, 0, 0, 0, 2, 2),
               booktabs = T, 
               col.names = c("Region", "Sex", "n", "a", "b", "c", "$R^2$", "Reference"),
               caption = "\\label{tab:all_params}Summary of 18 allometric growth parameters available for lionfish in the invaded range from peer-reviewed literature and this study. All parameters have been adjusted to convert from millimeters to grams. n = Sample size, Sex specifies whether data was presented for Females (F), Males (M), or both genders combined (B), a = scaling parameter for Eq. 1 (presented in $\\times 10^{-5}$), c = y-intercept for Eq. 3, b = exponent or slope for Eq. 1 or Eq. 3, respectively. The $R^2$ column indicates reported model fit.")
```

```{r, fig.height = 3.8, fig.pos = "H", fig.cap = "\\label{fig:all_allo}Length-weight relationships (n = 18) for 12 studies and this study. Colors indicate studies from which the parameters were extracted. Solid lines indicate that the fit was performed for males and females pooled together. Dotted lines indicate that the regression was performed on females, and dashed lines indicate it was performed for males. The dashed black line represents the relationship estimated in this study."}

review %>% 
  filter(!Study == "This study") %>% 
  mutate(Length = list(seq(1, 400))) %>% 
  unnest() %>% 
  mutate(Weight = a*(Length^b),
         This = 10^(coefficients(model)[1] + coefficients(model)[2]*log10(Length))) %>%
  group_by(Length) %>% 
  mutate(min = min(Weight),
         max = max(Weight)) %>% 
  ungroup() %>% 
  ggplot(aes(x = Length)) +
  geom_line(aes(y = Weight, color = Study, linetype = Gender, group = Study_id), size = 1) +
  geom_line(aes(y = This), size = 1.2, color = "black", linetype = "dashed") +
  labs(x = "Total length (mm)", y = "Total Weight (gr)") +
  theme_cowplot() +
  theme(text = element_text(family = "Times"),
        legend.justification=c(0,1),
        legend.position=c(0.01, 1),
        legend.background = element_rect(fill = "transparent"),
        legend.box = "horizontal",
        legend.text = element_text(size = 9),
        legend.title = element_text(size = 10)) +
  guides(color = guide_legend(order = 1),
         linetype = guide_legend(order = 0, ncol = 2),
         size = "none") +
  scale_color_brewer(palette = "Paired", direction = -1) +
  scale_linetype_manual(values=c("solid", "longdash", "dotted")) +
  scale_size_manual(values = c(1, 1.5))
```

```{r, fig.height = 4.2, fig.width = 5.5, fig.cap = "\\label{fig:bio_ratio}Violin plot of predicted to observed weight ratios for 18 pairs of allometric parameters. Red and blue circles indicate median and mean values, respectively. Like letters indicate values that do not differ significantly (Tukey's HSD; p < 0.05)."}
groups$groups%<>%
  mutate(Study_id = rownames(.)) %>% 
  select(-Biomass_ratio)

@@

We  detected substantial differences in weight-at-length between organisms from the Caribbean and Gulf of Mexico / North-Western Atlantic. Weight estimates using parameters from the Gulf of Mexico and North-Western Atlantic were higher on average than those from the Caribbean. The average predicted-to-observed weight ratios from these three regions were insert 1.24 $\pm$ 0.309, 1.76 $\pm$ 0.496, and Caribbean 1.17 $\pm$ 0.398, respectively. These length-weight differences mirror similar findings of regional variability in age-at-length relationships of lionfish across both their invaded and native regions \citep{pusack_2016}. These differences may be driven by genetic variation or by organisms being exposed to distinct environmental conditions. For example, \citet{betancurr_2011} used mitochondrial DNA to demonstrate the existence of two distinct population groups, identified as the “Caribbean group” and “Northern Group”, and \citet{fogg_2015} alternatively suggested that age-at-length differences may be driven by climate. Differences in weight-at-length could also reflect differences in energy input (*i.e.* in some regions, lionfish eat more) or differential usage of this energy (*e.g.* regional differences in predator abundances lead to different usage of energy), or a combination of both. Future research is needed to determine which processes are at work here.

Differences in length-weight relationships have traditionally been highlighted as potential pitfalls to fishery management. For example, \citet{wilson_2012} show that small-scale variations in length-at-age and fishing mortality in other Scorpaeniformes (*Sebastes rastrelliger*) translate to differential landings, effort, and catch per unit effort in the live fish fishery of California, and that these differences must be taken into account in management plans. The lionfish case poses the opposite scenario, where the manager desires to erradicate species. To accurately gauge both the effectiveness of lionfish removal efforts and the resources needed to successfully manage an invasion, we must acknowledge and understand regional biological differences in important variables such as allometric growth parameters.

>>>>

---
title: Invasive lionfish present region-wide variation of allometric growth in the Western Atlantic
abstract: "Lionfish (*Pterois volitans / miles*) are an invasive species in the Western Atlantic and the Caribbean. Improving management of invasive lionfish populations requires accurate estimates of total biomass, which depend on accurate estimates of the length-weight relationship. Lionfish exhibit a sedentary behavior that is likely to result in high levels of spatial variation in some life history characterstics, such as the length-weight relationship. We review 12 published length-weight relationships of lionfish taken throughout their invasive range and describe the spatial variation and implications of these differences. The spatial pattern observed here is consistent with those reported in genetic studies or age-at-length studies. We show that the use of *ex situ* parameters can result in up to a three-fold overestimation of Total Weight, but that using parameters obtained nearby reduces this error. These findings can have major implications for management in terms of predicting effects on local ecosystems, evaluating the effectiveness of removal programs, or estimating biomass available for harvest."
output:
  pdf_document:
    keep_tex: yes
    number_sections: no
    template: peerj_template.tex
---

<!-- Alternative titles: -->
<!-- Region-wide variation of allometric growth in invasive lionfish -->

@@

```{r load packages}
suppressPackageStartupMessages({
  library(stringr)
  library(agricolae)
  library(magrittr)
  library(stargazer)
  library(sandwich)
  library(lmtest)
  library(broom)
  library(tmap)
  library(rgdal)
  library(sf)
  library(cowplot)
  library(tidyverse)
})
```

@@

Lionfish (*Pterois volitans/miles* complex) are an invasive species in the western Atlantic and Caribbean Sea, likely introduced through liberation of aquarium-kept organisms \citep{betancurr_2011}. They are the first invasive marine vertebrates established along the North Atlantic Caribbean coasts \citep{schofield_2009,schofield_2010,sabidoitza_2016} and their presence has been labeled as a major marine invasion because they threaten local biodiversity, spread rapidly, and are difficult to manage \citep{hixon_2016}.

Invasive lionfish are primarily studied in coral reef ecosystems, where their impacts are far-reaching. For example, field experiments by \citet{albins_2008} showed that lionfish establishment led to reduced recruitment of native fishes by nearly 80% over a five week period in Florida. \citet{green_2012} report that prey fish biomass declined 65% over two years as lionfish biomass increased along Bahamian coral reefs. Lionfish have also established invasive populations in other habitats such as estuaries, mangroves, hard-bottomed areas, and mesophotic reefs \citep{jud_2011,barbour_2010,muoz_2011,andradibrown_2017,claydon_2012}.

A substantial amount of research describes lionfish feeding ecology throughout the invaded range \citep{muoz_2011,morris_2009,cote_2013,dahl_2014,valdezmoreno_2012,villaseorderbez_2014,hackerott_2017,sandel_2015}. A meta-analysis by  \citet{peake_2018} showed that invasive lionfish prey on at least 167 different species across the tropical and temperate North Atlantic. Their feeding behavior and high consumption rates can reduce recruitment and population sizes of native reef-fish species, and can further endanger reef fish \citep{albins_2008, green_2012,rocha_2015}. (However, see \citet{hackerott_2017} for a case where there was no evidence that lionfish affected the density, richness, or community composition of prey fishes). Major efforts have been made to understand the possible impacts of the invasion by tracking the spread of established lionfish populations through time \citep{schofield_2009,schofield_2010} and by predicting invasion ranges under future climates \citep{grieve_2016}. It{s been proposed that trophic impacts of lionfish can be minimized if local lionfish biomass is controled by culling \citep{ariasgonzalez_2011}.

@@

The feasibility of establishing fisheries through lionfish removal programs has been extensively evaluated through field observations and empirical modeling \citep{barbour_2011,morris_2011,deleon_2013,johnston_2015,sandel_2015,chin_2016,usseglio_2017}. One contributing factor to the success of many removal programs is the sedentary nature of adult lionfish \citep{jud_2012}. Culling programs are effective in reducing local adult populations largely because lionfish exhibit high levels of site fidelity and rarely leave their home range \citep{Fishelson_1997,cote_2014,kochzius_2005}. Fish with this sedentary behavior are likely to exhibit high levels of spatial variation in important life history characterstics such as growth or natural mortality rates \citep{hutchinson_2008,wilson_2012}. The importance of considering spatial heterogeneity is well-documented in terms of assessing and managing sedentary species \citep{gunderson_2008,guan_2013}, and such variation should be accounted for when evaluating the feasibility of establishing lionfish fisheries as well.

Determining the feasibility of such initiatives require modeling changes in biomass in response to changes in mortality (*i.e.* culling). A common way to model this is via length-structured population models, where fish lengths are converted to  weight to calculate total biomass \citep{cote_2014,barbour_2011,andradibrown_2017}. The length-weight relationship is therefore an essential component of these models, but this relationship can vary across regions as a response to biotic (*e.g.* local food availability) and abiotic (*e.g.* water temperature) conditions \citep{johnson_2016}. Literature suggests that site-specific parameters are necessary in order to accurately estimate biomass when length-weight relationships are spatially variable, and this variability becomes increasingly important when estimating the potential effectiveness and feasability of lionfish culling programs or when identifying total biomass available for harvest by fishers \citep{barbour_2011,morris_2011,johnston_2015,chin_2016,cote_2014}. In addition to environmentally-driven spatial variation, genetic analysis of invasive lionfish suggest biological differences due to the existence of two genetically distinct subpopulations between the northwest Atlantic and the Caribbean \citep{betancurr_2011}. To date, no studies have examined region-wide differences in length-weight parameters despite the large number of studies reporting this relationship for lionfish.

@@

We reviewed 12 published studies and obtained 17 length-weight relationships for the North Atlantic (n = 1), Gulf of Mexico (n = 7), and Caribbean (n = 10, Table \ref{tab:all_params}, Fig \ref{fig:all_allo}). We collected information on sampling methods, sex differentiation, location, and depth ranges from each study when available, and assumed both genders were included in a study if gender was unspecified. Only two studies reported parameters for each gender \citep{aguilarperera_2016,fogg_2013}. Reviewed studies presented information for organisms obtained at depths between `r min(review$minD, na.rm = T)` m and `r max(review$maxD, na.rm = T)` m. Three studies explicitly stated that their organisms were sampled with pole spears \citep{aguilarperera_2016,chin_2016,dahl_2014,sabidoitz_2016}, and five studies mentioned that some of their organisms were obtained with pole spears (or other type of harpoon) but also hand-held nets or fish traps \citep{sandel_2015,barbour_2011,fogg_2013,edwards_2014,sabidoitza_2016,sabidoitz_2016,toledohernndez_2014}, and two studies did not specify how organisms were sampled \citep{deleon_2013,darling_2011}. \citet{fogg_2013} use spine-less weight in the length-weight relationship estimation, and thus their parameters likely underestimate total wieght. Since no spine-less to total weight conversions were available, these parameters were taken as reported.

@@

```{r, fig.cap = "\\label{fig:map}Locations where allometric growth parameters of lionfish (*Pterois spp*) have been reported. Circle sizes indicate sample size from each study, colors indicate the $b$ coefficient from Eq. \\ref{eq:allometric}.", fig.height = 3.5}

@@

# Map with SF
ggplot() +
  geom_sf(data = W) +
  geom_point(data = review %>%
      mutate(N = ifelse(is.na(N), 1, N)),
    aes(x = lon, y = lat, fill = b, size = N), shape = 21) +
  theme_classic() +
  scale_fill_gradientn(colors = colorRamps::matlab.like(20)) +
  theme(legend.background = element_blank(),
        panel.grid.major = element_line(colour = "transparent")) +
  lims(x = c(-100, -60), y = c(9, 36)) +
  labs(x = "longitude", y = "latitude") +
  geom_text(aes(x = -70, y = 30), label = "North Atlantic\nOcean") +
  geom_text(aes(x = -90, y = 25), label = "Gulf of Mexico") +
  geom_text(aes(x = -75, y = 15), label = "Caribbean\nSea")

@@

Combining the length-weight parameters extracted from the literature and the additional pair calculated here, we obtain a total of 18 pairs. We use the central Mexican Caribbean as a case study of how the use of *ex situ* parameters influences the accuracy of weight estimates for lionfish. We estimated TW from the TL observations we collected in the central Mexican Caribbean (n = 109) using each of the 18 parameter pairs and divided predicted weights by known observed weights to obtain a simple measure of over- or underestimation. Difference in mean weight ratios across the different parameter pairs were tested with a one-way analysis of variance (ANOVA) and Tuskey's test was used for post-hoc tests. All analyses were performed in R version `r paste(version$major, version$minor, sep = ".")` \citep{rcore_2018}. Raw data and code used in this work are available at dryad.org.

@@

The length-weight relationship for organisms from the central Mexican Caribbean resulted in the coefficient values $a = `r 10^c`$, $b = `r b`$ and $c = `r c`$ ($R^2 = `r Rsq`$, `r Fstring`, $p < 0.001$). The allometric factor ($b$) was significantly different from $b = 3$ ($t(107) = `r formatC((coef(model)[[2]]-3)/robust_se[[2]], digits = 2, format = "f")`; p<0.001$) indicating that lionfish present allometric growth. The length-weight coefficients estimated in this study were within the range identified by studies in other regions (Table \ref{tab:all_params}). Figure \ref{fig:l-w-carib} shows the relationship between TL and TW for this region, and more information on model fit is presented in Table S2.

Figure \ref{fig:all_allo} shows the length-weight relationships with parameters from all studies. Parameters from models fit to males or females exclusively tend to have a higher steepness (*i.e.* higher allometric parameter), with mean $\pm$ standard deviation values of $b = `r formatC(mean(review$b[review$Gender == "Male"]), digits = 2, format = "f")` \pm `r formatC(sd(review$b[review$Gender == "Male"]), digits = 2, format = "f")`$ and $b = `r formatC(mean(review$b[review$Gender == "Female"]), digits = 2, format = "f")` \pm `r formatC(sd(review$b[review$Gender == "Female"]), digits = 2, format = "f")`$ for males and females respectively, compared to parameters from models for combined genders with a mean $\pm$ standard deviation value of $b = `r formatC(mean(review$b[review$Gender == "Both"]), digits = 2, format = "f")` \pm `r formatC(sd(review$b[review$Gender == "Both"]), digits = 2, format = "f")`$. In the case of the ponderal index ($a$) and its $log_{10}$ transformation ($c$), values were higher for parameters for combined genders.

There were significant differences in our predicted weights for the central Mexican Caribbean when using the different pairs of parameters ($F(df = 15; 1728) = 38.26; p < 0.001$). Weight estimates using parameters from the Gulf of Mexico and North-Western Atlantic were higher on average than those from the Caribbean. The average ($\pm$ SD) predicted-to-observed weight ratios from these three regions were 1.24 $\pm$ 0.309, 1.76 $\pm$ 0.496, and 1.17 $\pm$ 0.398, respectively. The lowest weight estimates resulted from using the allometric parameters from Banco Chinchorro in the Caribbean \citet{sabidoitz_2016}, and the highest weight estimates came from the Northern Atlantic \citep{barbour_2011}. The calculated ratio of predicted-to-observed weight  ranged from `r filter(Predicted_biomass, Study_id == "Sabido2BC") %$% mean(Biomass_ratio) %>% formatC(digits = 2, format = "f")` $\pm$ `r filter(Predicted_biomass, Study_id == "Sabido2BC") %$% sd(Biomass_ratio) %>% formatC(digits = 2, format = "f")` to `r filter(Predicted_biomass, Study_id == "Barbour") %$% mean(Biomass_ratio) %>% formatC(digits = 2, format = "f")` $\pm$ `r filter(Predicted_biomass, Study_id == "Barbour") %$% sd(Biomass_ratio) %>% formatC(digits = 2, format = "f")` (mean $\pm$ SD). Tukey's post-hoc test suggests that weight ratios for the central Mexican Caribbean were not differnt from those obntained with parameters from Little Cayman, the Bahamas, and some sites in the Gulf of Mexico (Tukeys HSD $p > 0.05$). Predicted-to-observed weight ratios are presented in Figure \ref{fig:bio_ratio}. Spine-less weight parameters from \citet{fogg_2013} still produced predicted-to-observed weight ratios > 1.

@@

```{r}
review %>% 
  select(Region, Gender2, N, a, b, c, r2, Study) %>% 
  filter(!Study == "Fishbase") %>% 
  arrange(Region) %>% 
  mutate(r2 = ifelse(is.na(r2), "-", r2),
         N = ifelse(is.na(N), "-", N),
         a = formatC((a*1e5), digits = 2)) %>% 
  knitr::kable(format = "latex",
               row.names = F,
               digits =  c(0, 0, 0, 0, 2, 2),
               booktabs = T,
               # longtable = T,
               col.names = c("Region", "Sex", "n", "a", "b", "c", "$R^2$", "Reference"),
               caption = "\\label{tab:all_params}Summary of 18 allometric growth parameters available for lionfish in the invaded range from peer-reviewed literature and this study. All parameters have been adjusted to convert from millimeters to grams. n = Sample size, Sex specifies whether data was presented for Females (F), Males (M), or both genders combined (B), a = scaling parameter for Eq. 1 (presented in $\\times 10^{-5}$), c = y-intercept for Eq. 3, b = exponent or slope for Eq. 1 or Eq. 3, respectively. The $R^2$ column indicates reported model fit.")
```

```{r, fig.height = 3.6, fig.pos = "H", fig.cap = "\\label{fig:all_allo}Length-weight relationships (n = 18) for 12 studies and this study. Colors indicate studies from which the parameters were extracted. Dotted, dashed and solid lines show models for males, females, and combined sexes, respectively. The dashed black line represents the relationship estimated in this study."}

review %>% 
  filter(!Study == "This study") %>% 
  mutate(Length = list(seq(1, 400))) %>% 
  unnest() %>% 
  mutate(Weight = a*(Length^b),
         This = 10^(coefficients(model)[1] + coefficients(model)[2]*log10(Length))) %>%
  group_by(Length) %>% 
  mutate(min = min(Weight),
         max = max(Weight)) %>% 
  ungroup() %>% 
  ggplot(aes(x = Length)) +
  geom_line(aes(y = Weight, color = Study, linetype = Gender, group = Study_id), size = 1) +
  geom_line(aes(y = This), size = 1.2, color = "black", linetype = "dashed") +
  labs(x = "Total length (mm)", y = "Total Weight (gr)") +
  theme_cowplot() +
  theme(text = element_text(family = "Times"),
        legend.justification=c(0,1),
        legend.position=c(0.01, 1.05),
        legend.background = element_rect(fill = "transparent"),
        legend.box = "horizontal",
        legend.text = element_text(size = 9),
        legend.title = element_text(size = 10)) +
  guides(color = guide_legend(order = 1),
         linetype = guide_legend(order = 0, ncol = 2),
         size = "none") +
  scale_color_brewer(palette = "Paired", direction = -1) +
  scale_linetype_manual(values=c("solid", "dashed", "dotted"), name = "Sex") +
  scale_size_manual(values = c(1, 1.5))
```

```{r, fig.height = 4, fig.width = 5.5, fig.cap = "\\label{fig:bio_ratio}Violin plot of predicted-to-observed weight ratios for 18 pairs of allometric parameters. Red and blue circles indicate median and mean values, respectively. Like letters indicate values that do not differ significantly."}
groups$groups%<>%
  mutate(Study_id = rownames(.)) %>% 
  select(-Biomass_ratio)

@@

We  detected substantial differences in weight-at-length between organisms from the Caribbean and Gulf of Mexico / North-Western Atlantic. The groupings of predicted-to-observed weight ratios were consistent with the spatial distribution of the studies, suggesting that these differences are mediated by space. These length-weight differences mirror similar findings of regional variability in age-at-length relationships of lionfish across both their invaded and native regions \citep{pusack_2016}. These patterns may be driven by genetic variation or by organisms being exposed to distinct environmental conditions. For example, \citet{betancurr_2011} used mitochondrial DNA to demonstrate the existence of two distinct population groups, identified as the “Caribbean group” and “Northern Group”, and \citet{fogg_2015} alternatively suggested that age-at-length differences may be driven by climate. Differences in weight-at-length could also reflect differences in energy input or differential usage of this energy, or a combination of both. Future research is needed to determine which processes are at work here.

Differences in length-weight relationships have traditionally been highlighted as potential pitfalls to fishery management. For example, \citet{wilson_2012} show that small-scale variations in length-at-age and fishing mortality in other Scorpaeniformes translate to differential landings, effort, and catch per unit effort in the live fish fishery of California, and that these differences must be taken into account in management plans. The lionfish case poses the opposite scenario, where the manager desires to erradicate species. To accurately gauge both the effectiveness of lionfish removal efforts and the resources needed to successfully manage an invasion, we must acknowledge and understand regional biological differences in important variables such as allometric growth parameters.

The results presented here have major implications for management. For example, \citet{edwards_2014} simulated a lionfish culling program under two scenarios, one using length-at-age and length-to-weight parameters from North Carolina and one using parameters from Little Cayman. Their results show that using different parameters caused up to a four-year difference in the time required for the simulated lionfish population to recover to 90% of its initial biomass after removals ceased. Here, we show that using one set of length-weight parameters versus another for a given length can result in up to a threefold TW overestimation. These spatially-driven differences become especially important when allocating resources for lionfish removal programs, incentivizing its fishery as a source of alternative livelihoods, or estimating ecosystem impacts. Research efforts focused on invasive lionfish populations need to use parameters calculated for their region to the extent possible, or at least use reasonable sets of different parameters that provide upper and lower bounds in their results.
